### 2010-10-11 pugal-esco2010.tex
Minor changes (address etc).

\documentclass[letterpaper,3p]{elsarticle}

\journal{Applied Mathematics and Computation}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{subfig}

\begin{document}
\begin{frontmatter}

\title{\emph{hp}-Finite Element Model of Poisson and Nernst-Planck\\ System of Equations}

\author[unrme,tartu]{D.~Pugal}
\ead{david.pugal@gmail.com}

\author[unrmath,czech]{P.~Solin\corref{cor2}}
\ead{solin@unr.edu}

\author[unrme]{K.~J.~Kim\corref{cor1}}
\ead{kwangkim@unr.edu}

\author[tartu]{A.~Aabloo}
\ead{alvo@ut.ee}

\address[unrme]{Mechanical Engineering Department, University of Nevada, Reno, NV, U.S.A.}
\address[unrmath]{Department of Mathematics and Statistics,
University of Nevada, Reno, NV, U.S.A.}
\address[czech]{Institute of Thermomechanics, Prague, Czech Republic}
\address[tartu]{Institute of Technology, Tartu University, Estonia}

\cortext[cor1]{Corresponding author}
\cortext[cor2]{Principal corresponding author}

\begin{abstract}
 In this paper we present a \emph{hp}-finite element
model (\emph{hp}-FEM) of Nernst-Planck and Poisson equation system.
The model is implemented in Hermes which is a
space- and space-time adaptive \emph{hp}-FEM solver. The time dependent
adaptivity is used to control the error of the solution. Full mathematical
derivation of the weak formulation of the system of equations and
the solution comparison with a nonadaptive conventional FEM is presented.
Furthermore, we extend the discussion on Hermes3D as a modeling tool that can
help in studying and improving the IPMC materials.
\end{abstract}

\begin{keyword}
  \emph{hp}-FEM \sep Nernst-Planck \sep Poisson \sep weak form
\end{keyword}

\end{frontmatter}

\input{intro}
\input{motivation}
\input{model}
\input{results}
\input{conclusions}

\section*{Acknowledgments}
KJK thanks the financial support from the Office of Naval Research under Award N000140910218.

\bibliographystyle{elsarticle-num}    % Bibliography: Author-Date system
\bibliography{pugal-esco2010}      % pls. call your database here

\end{document}