### 2010-10-19 pugal-esco2010.tex

\documentclass[mathpazo]{cicp}

%\journal{Applied Mathematics and Computation}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsrefs}
%\usepackage{subfig}

\begin{document}

%\title{\emph{hp}-Finite Element Model of Poisson and Nernst-Planck\\ System of Equations}
\title{Modeling Ionic Polymer-Metal Composites\\ with Space-Time Adaptive Multimesh \emph{hp}-FEM}

\author[Deivid Pugal et.~al.]{Deivid Pugal\affil{1}\comma\affil{2},
Pavel Solin\affil{3}\comma\affil{4}\comma\corrauth,
Kwang J. Kim\affil{1}, and Alvo Aabloo\affil{2}}

\address{\affilnum{1}\ Mechanical Engineering Department, University of Nevada, Reno, NV, U.S.A.\\
\affilnum{2}\ Institute of Technology, Tartu University, Estonia\\
\affilnum{3}\ Department of Mathematics and Statistics, University of Nevada, Reno, NV, U.S.A.\\
\affilnum{4}\ Institute of Thermomechanics, Prague, Czech Republic}

\emails{{\tt david.pugal@gmail.com} (D.~Pugal), {\tt solin@unr.edu} (P.~Solin),
	{\tt kwangkim@unr.edu} (K.~Kim), {\tt alvo@ut.ee} (A.~Aabloo)}

%\author[unrme,tartu]{D.~Pugal}
%\ead{david.pugal@gmail.com}
%
%\author[unrmath,czech]{P.~Solin\corref{cor2}}
%\ead{solin@unr.edu}
%
%\author[unrme]{K.~J.~Kim\corref{cor1}}
%\ead{kwangkim@unr.edu}
%
%\author[tartu]{A.~Aabloo}
%\ead{alvo@ut.ee}
%
%\address[unrme]{Mechanical Engineering Department, University of Nevada, Reno, NV, U.S.A.}
%\address[unrmath]{Department of Mathematics and Statistics,
%University of Nevada, Reno, NV, U.S.A.}
%\address[czech]{Institute of Thermomechanics, Prague, Czech Republic}
%\address[tartu]{Institute of Technology, Tartu University, Estonia}
%
%\cortext[cor1]{Corresponding author}
%\cortext[cor2]{Principal corresponding author}

\begin{abstract}
In this paper we study a multiphysics coupled problem consisting 
of the Poisson and Nernst-Planck equations, that is used to 
model ionic polymer-metal composite (IPMC) materials. 
The problem is discretized using the finite element method (FEM) and solved using 
space-time adaptive \emph{hp}-FEM algorithms provided by the Hermes 
library. Numerical results are presented and discussed, including 
comparison of various versions of adaptive $hp$-FEM and conventional 
low-order FEM. In particular, we discuss the performance of a novel
adaptive multimesh $hp$-FEM algorithm.
\end{abstract}

%\begin{keyword}
%\keywords{Ionic polymer-metal composites, IPMC, Nernst-Planck equation, Poisson equation,
%finite element method, space-time adaptivity, multimesh \emph{hp}-FEM }
%\end{keyword}

%see http://www.ams.org/mathscinet/msc/msc2010.html?t=35Qxx&btn=Current

\ams{35Q84, 35Q05, 35J47}
\keywords{Ionic polymer-metal composites, IPMC, Nernst-Planck equation, Poisson equation,
finite element method, space-time adaptivity, multimesh \emph{hp}-FEM }
\maketitle

%\end{frontmatter}

\input{intro}
\input{motivation}
\input{model}
\input{results}
\input{conclusions}

\section*{Acknowledgments}
The second author was partially supported by the Grant Agency of the Academy of 
Sciences of the Czech Republic under Grant No. IAA100760702, and by the U.S.
Department of Energy Research Subcontract No. 00089911. 
The third author acknowledges the financial support of the U.S. Office of Naval Research 
under Award N000140910218.

%\begin{thebibliography}{99}
%\bibitem{\ldots}..

%\end{thebibliography}

%\bibliographystyle{amsrs}    % Bibliography: Author-Date system
\bibliography{pugal-esco2010}      % pls. call your database here

\end{document}