### 2011-03-22 pugal-cicp2010.tex
Submitted final version

\documentclass[mathpazo]{cicp}

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsrefs}

\begin{document}

\title{Modeling Ionic Polymer-Metal Composites\\ with Space-Time Adaptive Multimesh \emph{hp}-FEM}

\author[Deivid Pugal et.~al.]{Deivid Pugal\affil{1}\comma\affil{4},
Pavel Solin\affil{2}\comma\affil{3}\comma\corrauth,
Kwang J. Kim\affil{1}, and Alvo Aabloo\affil{4}}

\address{\affilnum{1}\ Mechanical Engineering Department, University of Nevada, Reno, NV, U.S.A.\\
\affilnum{2}\ Department of Mathematics and Statistics, University of Nevada, Reno, NV, U.S.A.\\
\affilnum{3}\ Institute of Thermomechanics, Prague, Czech Republic\\
\affilnum{4}\ Institute of Technology, Tartu University, Estonia}

\emails{{\tt david.pugal@gmail.com} (D.~Pugal), {\tt solin@unr.edu} (P.~Solin),
	{\tt kwangkim@unr.edu} (K.~Kim), {\tt alvo@ut.ee} (A.~Aabloo)}

\begin{abstract}
We are concerned with a model of ionic polymer-metal composite (IPMC) materials
that consists of a coupled system of the Poisson and Nernst-Planck equations, 
discretized by means of the finite element method (FEM). We show that due to the 
transient character of the problem it is efficient to use adaptive algorithms 
that are capable of changing the mesh dynamically in time. We also show 
that due to large qualitative and quantitative differences between the 
two solution components, it is efficient to approximate them on different 
meshes using a novel adaptive multimesh \emph{hp}-FEM. The study is 
accompanied with numerous computations and comparisons of the adaptive 
multimesh \emph{hp}-FEM with several other adaptive FEM algorithms. 
\end{abstract}

\ams{35Q84, 35Q05, 35J47}
\keywords{Ionic polymer-metal composites, IPMC, Nernst-Planck equation, Poisson equation,
finite element method, FEM, adaptive multi-mesh \emph{hp}-FEM }
\maketitle

\input{intro}
\input{motivation}
\input{model}
\input{results}
\input{conclusions}

\section*{Acknowledgments}
The second author was partially supported by the Grant Agency of the Academy of 
Sciences of the Czech Republic under Grant No. IAA100760702, and by the U.S.
Department of Energy Research Subcontract No. 00089911. 
The third author acknowledges the financial support of the U.S. Office of Naval Research 
under Award N000140910218. The fourth author acknowledges the financial support of
the Estonian Ministry of Education, grant \#SF0180008s08.

\newpage

\bibliographystyle{amsrn}    % Bibliography: Author-Date system
%\bibliography{pugal-cicp2010}      % pls. call your database here
\begin{bibdiv}
\begin{biblist}

\bib{basu1997membrane}{article}{
      author={Basu, S.},
      author={Sharma, M.M.},
       title={{An improved space-charge model for flow through charged
  microporous membranes}},
        date={1997},
     journal={Journal of Membrane Science},
      volume={124},
      number={1},
       pages={77\ndash 91},
}

\bib{bazant2004diffuse}{article}{
      author={Bazant, M.Z.},
      author={Thornton, K.},
      author={Ajdari, A.},
       title={{Diffuse-charge dynamics in electrochemical systems}},
        date={2004},
        ISSN={1550-2376},
     journal={Physical Review E},
      volume={70},
      number={2},
       pages={21506},
}

\bib{dubcova2010space}{article}{
      author={Dubcova, L.},
      author={Solin, P.},
      author={Cerveny, J.},
      author={Kus, P.},
       title={{Space and Time Adaptive Two-Mesh $hp$-Finite Element Method for
  Transient Microwave Heating Problems}},
        date={2010},
     journal={Electromagnetics},
      volume={30},
      number={1},
       pages={23\ndash 40},
}

\bib{nasser2002applied}{article}{
      author={Nemat-Nasser, S.},
       title={Micromechanics of actuation of ionic polymer-metal composites},
        date={2002},
     journal={Journal of Applied Physics},
      volume={92},
      number={5},
       pages={2899\ndash 2915},
}

\bib{newbury2003intelligent}{article}{
      author={Newbury, K.M.},
      author={Leo, D.J.},
       title={{Linear electromechanical model of ionic polymer transducers-Part
  I: Model Development}},
        date={2003},
     journal={Journal of Intelligent Material Systems and Structures},
      volume={14},
      number={6},
       pages={333},
}

\bib{pugal2009}{inproceedings}{
      author={Pugal, D.},
      author={Aabloo, A.},
      author={Kim, K.~J.},
       title={Dynamic surface resistance model of ipmc},
        date={2009},
      editor={Ounaies, Zoubeida},
      editor={Li, Jiangyu},
      volume={7289},
   publisher={SPIE},
       pages={72891E},
}

\bib{pugal2010polymer}{article}{
      author={Pugal, D.},
      author={Jung, K.},
      author={Aabloo, A.},
      author={Kim, K.J.},
       title={{Ionic polymer-metal composite mechanoelectrical transduction:
  review and perspectives}},
        date={2010},
     journal={Polymer international},
      volume={59},
      number={3},
       pages={279\ndash 289},
}

\bib{pugal2010spie3d}{inproceedings}{
      author={Pugal, D.},
      author={Kim, K.~J.},
      author={Aabloo, A.},
       title={Modeling the transduction of ipmc in 3d configurations},
        date={2010},
      editor={Ounaies, Zoubeida},
      editor={Li, Jiangyu},
      volume={7644},
   publisher={SPIE},
       pages={76441T},
}

\bib{pugal2008appliedphysics}{article}{
      author={Pugal, D.},
      author={Kim, K.~J.},
      author={Punning, A.},
      author={Kasemagi, H.},
      author={Kruusmaa, M.},
      author={Aabloo, A.},
       title={A self-oscillating ionic polymer-metal composite bending
  actuator},
        date={2008},
     journal={Journal of Applied Physics},
      volume={103},
      number={8},
       pages={084908},
}

\bib{shahinpoor2001smartmat}{article}{
      author={Shahinpoor, M.},
      author={Kim, K.~J.},
       title={Ionic polymer-metal composites: I. fundamentals},
        date={2001},
     journal={Smart Materials and Structures},
      volume={10},
      number={4},
       pages={819},
}

\bib{solin2010pde}{article}{
      author={Solin, P.},
      author={Andrs, D.},
      author={Cerveny, J.},
      author={Simko, M.},
       title={{PDE-independent adaptive $hp$-FEM based on hierarchic extension
  of finite element spaces}},
        date={2010},
     journal={Journal of computational and applied mathematics},
      volume={233},
      number={12},
       pages={3086\ndash 3094},
}

\bib{solin2010monolithic}{article}{
      author={Solin, P.},
      author={Cerveny, J.},
      author={Dubcova, L.},
      author={Andrs, D.},
       title={{Monolithic discretization of linear thermoelasticity problems
  via adaptive multimesh hp-FEM}},
        date={2010},
     journal={Journal of computational and applied mathematics},
      volume={234},
      number={7},
       pages={2350\ndash 2357},
}

\bib{solini2010adaptive}{article}{
      author={Solin, P.},
      author={Dubcova, L.},
      author={Dolezel, I.},
       title={{Adaptive $hp$-FEM with arbitrary-level hanging nodes for
  Maxwell’s equations}},
        date={2010},
     journal={Advances in Applied Mathematics and Mechanics},
      volume={2},
      number={4},
       pages={518\ndash 532},
}

\bib{solin2010adaptive}{article}{
      author={Solin, P.},
      author={Dubcova, L.},
      author={Kruis, J.},
       title={{Adaptive $hp$-FEM with dynamical meshes for transient heat and
  moisture transfer problems}},
        date={2010},
     journal={Journal of computational and applied mathematics},
      volume={233},
      number={12},
       pages={3103\ndash 3112},
}

\bib{Hermes-book}{book}{
      author={Solin, P.},
      author={Segeth, K.},
      author={Dolezel, I.},
       title={Higher-order finite element methods},
   publisher={Chapman \& Hall / CRC Press},
        date={2003},
}

\bib{valli2002control}{article}{
      author={Valli, A. M.~P.},
      author={Carey, G.~F.},
      author={Coutinho, A.},
       title={{Control strategies for timestep selection in simulation of
  coupled viscous flow and heat transfer}},
        date={2002},
     journal={Communications in Numerical Methods in Engineering},
      volume={18},
      number={2},
       pages={131\ndash 139},
}

\bib{wallmersperger2007appliedphysics}{article}{
      author={Wallmersperger, T.},
      author={Leo, D.~J.},
      author={Kothera, C.~S.},
       title={Transport modeling in ionomeric polymer transducers and its
  relationship to electromechanical coupling},
        date={2007},
     journal={Journal of Applied Physics},
      volume={101},
      number={2},
       pages={024912},
}

\end{biblist}
\end{bibdiv}

\end{document}